Learning to Synthesize Rules for Network Intrusion Detectors

### 2020-02-22 paper/paper.tex
examples

In the following, we discuss the rule produced by \tname{} to capture
this attack. The action and the header of the rules that \tname{}
creates are either inferred or left constant. For example, in this
case, \tname{} sets the action to \CodeIn{alert} and sets the header
to \CodeIn{<proto> any any -> any any}, indicating that it infers the
target protocol and uses the most general flow description possible,
\ie{}, it instructs the \nids\ to analyze the traffic flowing from/to
any address and port.  We understand that the choice of these
parameters are domain-specific and assume that the sys admin would
know how (and want) to customize them.

The main challenge in synthesizing rules is to discover the set of
options in a rule that blocks the malicious traffic, but allows the
benign traffic to pass. \tname{} proceeds as follows to synthesize
rules for this attack. First, \tname{} analyzes the malicious traffic and extracts options
expressed in that traffic. Recall from
Section~\ref{sec:rules-and-packets} that rule options can be inferred
from packet contents. In this case, however, that information alone is
insufficient as the malicious traffic is characterized as a sequence
of messages. \tname{} finds the following set of options at this stage by
analyzing the flow and contents of the messages:

\noindent
By construction, a rule that uses this set of options captures the
malicious traffic. Unfortunately, this rule is potentially
overspecified. Informally, the rule includes ``too specific''
characteristics of the malicious traffic and could potentially result
in false negatives. For example, in this particular case, note that
the option \CodeIn{window: 512} is not included in the golden rule
from Figure~\ref{fig:synflood-example}.

Second, \tname{} uses benign traffic to mitigate the overfitting
issue. In this step, \tname{} searches for alternative rules that
preserves the invarianxt that a rule should capture the positive
malicious traffic but not capture the negative benign traffic. It uses
the rule obtained in the previous step, containing three options, to
bootstrap the search for alternative rules. The search generates rules
containing subsets of the options from the seed rule.

%% \Luc{Ataque "JBoss JMX Console Beanshell Deployer WAR Upload and Deployment Exploit"
%% . O payload do pacote é o seguinte:
%% \begin{figure}[H]
%%   \lstinputlisting[language=C,numbers=none,keywords={dsize,itype}]{adaptor-rule.suricata}
%%   \label{fig:rule-example}
%% \end{figure}

%% A partir disso produzimos uma regra com os seguintes contents:

%% \begin{figure}[H]
%%   \lstinputlisting[language=C,numbers=none,keywords={dsize,itype}]{adaptor-contents-splitted.suricata}
%%   \label{fig:rule-example}
%% \end{figure}

%% Depois da etapa da minimização, as top 5 regras são:

%% \begin{figure}[H]
%%   \lstinputlisting[language=C,numbers=none,keywords={dsize,itype}]{adaptor-top5-rules.suricata}
%%   \label{fig:rule-example}
%% \end{figure}

%% A regra de ouro é:

%% \begin{figure}[H]
%%   \lstinputlisting[language=C,numbers=none,keywords={dsize,itype}]{adaptor-golden-rule.suricata}
%%   \label{fig:rule-example}
%% \end{figure}

Finally, \tname{} ranks the inferred rules according to heuristics
learned from a database of existing rules. For example, one heuristic
function \tname{} uses for ranking leverages the amount of options a
rule of a given protocol typically has---the score of a rule on this
function is proportional to the popularity of the number of options it
declares. The list below shows the rule options for the top-5 rules
\tname{} produces for this attack.\Mar{algum problema em fazer isto?
  exemplos concretos ajudam muito!} \Luc{No caso desse ataque, a regra inicial possui apenas duas opcoes (sem contar o threshold). Entao existem apenas 3 combinacoes possiveis, dessas 3 apenas 2 sao capazes de isolar o ataque do trafego normal. Portanto, o top-2 eh:
  \begin{figure}[h]
    \vspace{-2ex}
    \lstinputlisting[language=C,numbers=none,frame=none,keywords={window,flags}]{top-synflood.rules}
    \vspace{-3ex}  
  \end{figure}

To sum up, \tname{} uses the messages that manifest an attack in the
first step to create an overspecific seed rule, it uses the benign
traffic in the second step to create a set of plausible rules from the
seed rule, and it uses existing rules from popular sources in the
third step to rank alternative plausible rules.

@@

\vspace{-2ex}
\begin{figure}[h!]
  \lstinputlisting[language=C,numbers=none,keywords={dsize,itype}]{pingscan.suricata}
  \vspace{-2ex}    
  \caption{Suricata rule for Ping Scan attack.}
  \label{fig:pingscan-example}
\end{figure}
\vspace{-4ex}

Figure 4 shows a Suricata rule that captures a Ping Scan attempt, as
documented in NMAP~\cite{netmap}, a network discovery and security
auditing open-source tool. The relevant options are in bold---the
option \CodeIn{dsize: 0} indicates that the packet has no bytes in the
payload and the option \CodeIn{itype: 8} indicates that the packet to
be captured are ICMP Echo requests. Again, to synthesize rules for a
Ping Scan, \tname{} analyzes the positive malicious traffic and
extracts the following rule options:

\begin{figure}[h]
  \vspace{-2ex}
  \lstinputlisting[language=C,numbers=none,frame=none,keywords={dsize,icode,itype,icmp\_id,icmp\_seq}]{pingscan.suricata.synth}
  \vspace{-3ex}  
\end{figure}

\noindent
Note that, although options \CodeIn{dsize:0} and \CodeIn{itype:8} are
present in the set of inferred options, the others are not.  This
indicates, as expected, that the set of rule options is
overspecified. \tname{} produces the following five rules at the top
of the ranking after minimization and ranking.\Mar{eu ainda estou
  aguardando isto. A tabela 3 tambem nao indica o resultado!}

\begin{figure}[h]
  \vspace{-2ex}
  \lstinputlisting[language=C,numbers=none,frame=none,keywords={dsize,icode,itype,icmp\_id,icmp\_seq}]{top-pingscan.rules}
  \vspace{-3ex}  
\end{figure}

\Fix{...listem 5 primeiras regras neste caso...}

@@

>>>>

@@

\vspace{-1ex}
\begin{figure}[h!]
  \lstinputlisting[language=C,numbers=none,keywords={dsize,itype}]{pingscan.suricata}
  \vspace{-2ex}  
  \caption{Suricata rule for Ping Scan attack.}
  \label{fig:pingscan-example}
\end{figure}
\vspace{-3ex}

Figure 4 shows a Suricata rule that captures a Ping Scan attempt, as
documented in NMAP~\cite{netmap}, a network discovery and security
auditing open-source tool. The relevant options are in bold---the
option \CodeIn{dsize: 0} indicates that the packet has no bytes in the
payload and the option \CodeIn{itype: 8} indicates that the packet to
be captured are ICMP Echo requests.

In the following, we discuss the rule produced by \tname{} to capture
this attack. The action and the header of the rules that \tname{}
creates are either inferred or left constant. For example, in this
case, \tname{} sets the action to \CodeIn{alert} and sets the header
to \CodeIn{<proto> any any -> any any}, indicating that it infers the
target protocol and uses the most general flow description possible,
\ie{}, it instructs the \nids\ to analyze the traffic flowing from/to
any address and port.  The choice of these parameters are
domain-specific and subject to change by the system administrator.
The main challenge in synthesizing rules is to discover the set of
options in a rule that blocks the malicious traffic, but allows the
benign traffic to pass.

\tname{} proceeds as follows to synthesize rules. First, it analyzes the malicious
traffic and extracts options expressed in that traffic. Recall from
Section~\ref{sec:rules-and-packets} that rule options can be inferred
from packet contents. \tname{} finds the following set of options at
this stage:

\begin{figure}[h]
  \vspace{-2ex}
  \lstinputlisting[language=C,numbers=none,frame=none,keywords={dsize,icode,itype,icmp\_id,icmp\_seq}]{pingscan.suricata.synth}
  \vspace{-2ex}  
\end{figure}

\noindent
By construction, a rule that uses this set of options captures the
malicious traffic. Unfortunately, the rule above is
overspecified. Informally, the rule includes ``too specific''
characteristics of the malicious traffic and could potentially result
in false negatives. In this case, although options \CodeIn{dsize:0}
and \CodeIn{itype:8}~(see Figure~\ref{fig:pingscan-example}) are
present in the set of inferred options, the others are not.

Second, \tname{} uses the benign traffic to mitigate the overfitting
issue. In this step, \tname{} searches for alternative rules that
preserve the invariant to capture the positive malicious traffic but
not capture the negative benign traffic. It uses the rule obtained in
the previous step as seed to bootstrap the search for alternative
plausible rules, \ie{}, rules that satisfy the aforementioned
invariant. The search is designed to generate rules containing subsets
of the options from the seed rule. Considering this example, \tname{}
produces a total of \Fix{XXXXX} plausible rules at the end of the
search.

Finally, \tname{} ranks plausible rules according to heuristics
learned from a database of existing rules. The intuition is that new
rules should share characteristics with pre-existing rules. For
example, one heuristic function used by \tname{} leverages the amount
of options a rule typically has. The score of a rule, given by this
function, is proportional to the popularity of the number of options
declared in rules of a given protocol. At the end of the process,
\tname{} ranks the golden rule at the top of the ranking including
\Fix{XXXXXX} other plausible rules.

%% the following rules 
%% at the top of the ranking produced by \tname{}:
%% 1)~{\small\CodeIn{itype:8; dsize:0;}}, 2)~{\small\CodeIn{itype:8;
%%     icmp\_id:23570;}}, 3)~{\small\CodeIn{icode:8; icmp\_id:23570;}},
%% 4)~{\small\CodeIn{itype:8; icode:0; dsize:0;}},
%% 5)~{\small\CodeIn{icmp\_id:23570}}. Note that the golden rule appears
%% at the top position.

%% \begin{figure}[h]
%%   \vspace{-2ex}
%%   \lstinputlisting[language=C,numbers=none,frame=none,keywords={dsize,icode,itype,icmp\_id,icmp\_seq}]{top-pingscan.rules}
%%   \vspace{-3ex}  
%% \end{figure}

To sum up, \tname{} uses the messages that manifest an attack in the
first step to create a potentially overspecified seed rule, it uses
the benign traffic in the second step to create a set of plausible
rules from the seed rule, and it uses public databases with existing
rules in the third step to rank plausible rules.

@@

...........
In this case, however, that information alone is
insufficient as the malicious traffic is characterized as a sequence
of messages. 
...........

\noindent

.....

 The list below shows the rule options for the top-5 rules
\tname{} produces for this attack.\Mar{algum problema em fazer isto?
  exemplos concretos ajudam muito!} \Luc{No caso desse ataque, a regra inicial possui apenas duas opcoes (sem contar o threshold). Entao existem apenas 3 combinacoes possiveis, dessas 3 apenas 2 sao capazes de isolar o ataque do trafego normal. Portanto, o top-2 eh:
  \begin{figure}[h]
    \vspace{-2ex}
    \lstinputlisting[language=C,numbers=none,frame=none,keywords={window,flags}]{top-synflood.rules}
    \vspace{-3ex}  
  \end{figure}
